---
title: "Predicting how well an exercise is performed"
author: "Jordan Strauss"
date: "April 24, 2016"
output: html_document
---


# Summary
We will try to determine how well people perform a weightlifting excersize (dumbell curls) using data from accelerometers attached at different locations to their bodies


Load library's set seed
```{r}
library(caret);library(randomForest)
set.seed(12345)
```

#Data import
Download the dataset from the web and read into R.  Read blank entries as NAs
```{r}
download.file("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv", "MLCTrain.csv")          
TrainMaster<- read.csv("MLCTrain.csv", header=T, na.strings=c("","NA"))
```

# Exploratory Data Analysis

Check out the dataset by using summary and examining the class of the columns


```{r}
#summary(TrainMaster)
#lapply(TrainMaster,Class)
```

The response field is classe and it is spuriously correlated to the index field x.  It is also clear that there are a number of columns which are almost entirely NAs, except when the "new_window" column is equal to yes.  And there are a number of columns such as timestamps, user names and entry numbers which are not related to movements and should not be considered when training our models.  

#Cleaning the data

Filter the data such that the dataset only includes the predictor variables we want and the response

```{r}

#Filter out new window =yes
TrainMaster<-TrainMaster[TrainMaster$new_window!="yes",]
#Filter out Columns consisting of all NAs
TrainMaster <- TrainMaster[,colSums(is.na(TrainMaster))<nrow(TrainMaster)]
#Filter out Columns which are not the response or accellerometer data
TrainMaster<-TrainMaster[,-c(1,3,4,5,6,7)]
```


#Build Model
We split the data set into a training and a test set.  Since this is a classification problem and we the problem is straightfoward enough such that interpretability is not a major concern, we fit a model using Random Forest
```{r}
intrain<-createDataPartition(TrainMaster$classe,p=.7,list = FALSE)
training<-TrainMaster[intrain,]
testing<-TrainMaster[-intrain,]
modFit1<-randomForest(classe~.,data=training,importance=TRUE)
modFit1
```

Random forest fits a model with 500 trees,and an expected out of sample error of ~.5%

Plot variable importance
```{r}
varImpPlot(modFit1)
```


#Cross Validation
Predict values on the testing set and check accuracy
```{r}
pred1<-predict(modFit1,testing)
confusionMatrix(pred1,testing$classe)
```

Model shows an accuary of >99%

# Course Project Prediction Quiz

Download and import test data, predict classe using random forest model

```{r}
download.file("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv", "MLCTest.csv")          
test<-read.csv("MLCTest.csv",header = T,na.strings=c("","NA"))
predict(modFit1,test)
```
